### 2017-03-22 paper.tex
first compile

\documentclass[10pt,twocolumn]{article}
\usepackage{times}

% do not change these values
\baselineskip 12pt
\textheight 9in
\textwidth 6.5in
\oddsidemargin 0in
\topmargin 0in
\headheight 0in
\headsep 0in

\begin{document}

\title{ClaTex: LaTeX as a Service on the Cloud}
\author{John Doe$^1$ and Jane Monroe$^2$ \\
\small {\em  $^1$Dept. Cloud, Cloud University \quad
          $^2$Cloud National Labs} \\ [2mm]
\small Submission Type: Research
}
\date{}
\maketitle

\begin{abstract}
The time has come to offer LaTeX on the cloud.
\end{abstract}

\section{Introduction}

\section{Motivation}

\section{Design}

\section{Conclusion}

\bibliographystyle{abbrv}
\bibliography{biblio}

\end{document}